---
title: "Reproducible Research: Peer Assessment 1"
author: "jlaurensa"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r , echo = TRUE}
setwd("~/Coursera/Data_Science/Reproducible_Research/Peer_Assign_1/master")
Dataset <- read.table("activity.csv", header = TRUE, sep = ",", dec = ",")
```

## What is mean total number of steps taken per day?  
### 1. Make a histogram of the total number of steps taken each day
```{r Plot1_Histogram, echo =TRUE, result="asis"}
Dataset$date <- as.Date(Dataset$date)

par(mfrow = c(1,2)) 
plot(Dataset$date, Dataset$steps, type = "h", col = "blue", lwd = "2",
             main = " Barplot steps each day", xlab = "Date", ylab = "Frequency" )
hist(Dataset$steps, breaks = 100, col = "magenta", main = " Histogram steps each day", xlab = "Number of steps", ylab = "Frequency")
```

### 2. Calculate and report the mean and median of steps taken per day
```{r echo =TRUE, result="asis"}
library(plyr)
meanmedian <- ddply(Dataset,.(date), summarize, mean=mean(steps), median=median(steps))
meanmedian
```

## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r Plot2_Line, echo =TRUE, result="asis"}

Dataset$date <- as.Date(Dataset$date)
library(plyr)
Avg_Daily_Act <- ddply(Dataset,.(interval), summarize, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))

plot(Avg_Daily_Act$interval, Avg_Daily_Act$mean, type = "l", col = "blue", lwd = "1",
             main = " Total number of steps taken each day", xlab = "5-minute interval", ylab = "Avg steps taken (all days)")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo =TRUE, result="asis"}
SortAvg_Daily_Act <- arrange(Avg_Daily_Act, desc(mean))
head(SortAvg_Daily_Act)
max(Avg_Daily_Act$mean)
```

- According to the previous table, the 835th minute interval, on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo =TRUE, result="asis"}
bad <- is.na(Dataset)
miss_val <- nrow(Dataset[bad,])
```

- The total number of missing values in the dataset is `r miss_val`

### 2. Devise a strategy for filling in all of the missing values in the dataset. 
-  Strategy: The missing values in the dataset, [Steps variable], will be filled With 5-minute interval average across all the days in the dataset. Using `Avg_Daily_Act` table previously loaded.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in (variable NEWSTEPS).
```{r echo =TRUE, result="asis"}
newsteps <- vector()
for (i in 1:length(Dataset$steps)){
if (is.na(Dataset$steps[i]) == TRUE){
        y <- Dataset[i,3]
        y <- as.numeric(y)
        newsteps[i] <- round(subset(Avg_Daily_Act, interval == y)[,2],2)
} else { 
        newsteps[i] <- Dataset$steps[i]
        }
}
Dataset$newsteps <- newsteps
head(Dataset, 50)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r Plot3_Histogram, echo =TRUE, result="asis"}
Dataset$date <- as.Date(Dataset$date)

par(mfrow = c(1,2)) 
plot(Dataset$date, Dataset$newsteps, type = "h", col = "blue", lwd = "2",
             main = " Barplot steps each day", xlab = "Date", ylab = "Frequency" )
hist(Dataset$newsteps, breaks = 100, col = "magenta", main = " Filled data New steps", xlab = "New Number of steps", ylab = "Frequency")

library(plyr)
newdata <- ddply(Dataset,.(date), summarize, mean=mean(newsteps), median=median(newsteps))
cbind(newdata, meanmedian$mean, meanmedian$median)
```

- The impact of imputing missing data on the estimates of the total daily number of steps, is non existent, due to the strategy applied for filling the missing values.

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
### 2. panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r Plot4_Weekdayline, echo =TRUE, result="asis"}
library(lattice)
Dataset$weekday <- weekdays(Dataset$date)
Dataset$isweekday <- Dataset$weekday %in% c('sábado','domingo')
f <- factor(Dataset$isweekday, labels = c("Weekday", "Weekend"))
xyplot(Dataset$newsteps ~ Dataset$interval | f, type = "l", layout = c(1,2), xlab = "Interval", ylab = "Number of steps")
```

